---
title: "Ass2_Quant"
author: "Sage Grey"
date: "9/14/2020"
output: html_document
---

##### Downloading Libraries and Showing Variables:
```{r}
 library(tidycensus)
library(tidyverse)
library(ggplot2)
```

####  Load Household Data Set & Housekeeping
```{r}
# Load CSV
household_data <- read_csv("assignment1_vars.csv")
#Remove Scientific Notation
options(scipen = 999)
```

I Realized that I wanted to filter my data set to remove negative values for income, so instead of loading the csv, I am loading variables using the get_pums method. Came to this conclusion that I should filter for non negative values when talking with Emma Colley. I did this in the code from Assignment 1 so it is not shown in this R Markdown document. 


##Pulling Summary
```{r}
HINCP_summary <- summary(household_data$HINCP)

HINCP_summary
HINCP_summary["Max."]
```
## Another Way of Pulling Summary
```{r}
mean(household_data$HINCP)
quantile(household_data$HINCP)
sd(household_data$HINCP)
```

## 95=percent confidence interval
```{r}
confidence_int <-t.test(household_data$HINCP)
confidence_int

## Upper and Lower Limits
confidence_int$conf.int[1]
confidence_int$conf.int[2]
```

## Historgram 
```{r}
hist(household_data$HINCP)
```
## Histogram in GGPLOT
```{r}
ggplot(household_data, aes(x=HINCP)) +
  geom_histogram(color="green", fill ="purple", alpha =.5)
```
##Categorical Variable
```{r}
#Race
unique(household_data$RAC1P_label)
#Hispanic
unique(household_data$HISP_label)
```

