---
title: "Assignment 1"
author: "SVoorhees"
date: "9/4/2020"
output: html_document
---
```{r}
 library(tidycensus)
library(tidyverse)

```

## Overview & Goals
While compiling this dataset, I had in mind the question of how New Mexico is developing, specifically how New Mexico is building housing. New Mexico--unlike its neighbor Arizona--has seen relatively modest urban development in the last few decades, while maintaining a relatively stagnant overall population. However, there is nonetheless considerable development happening. In this course I would love to use the assignments to explore this development and provide insight into the questions of where is New Mexico developing, who is inhabiting these new developments, and what notable infrastructure/jobs are in these areas. 

## Variables & Unites Used
To answer the above question I have chosen to look at census track data using the ACS-5 Year Summary because it provides more direct information about construction. 

1. Household Income: **FHINCP** Categorical
2. When Structure was First Built: **YBL**  Continuous (mutated into)
3. Majority Race: **RAC1P** Categorical
4. First Mortgage Payment: **MRGP** Continuous    
5. Travel Time To Work: **JWMNP**


##Questions: 

If I wanted to compare the additional housing units. For Example, housing units between 2010 and 2000 census, how would I do this? 


```{r}
assignment1_vars <- get_pums(variables= c("YBL","FHINCP","RAC1P", "JWMNP", "MRGP"),
state= "NM",
year =2018,
survey ="acs1",
recode=TRUE) %>%
filter()%>%
mutate(age_struct = case_when(
    YBL_label == "1939 or earlier" ~ 100,
    YBL_label == "1940 to 1949" ~ 73,
    YBL_label == "1950 to 1959" ~ 63,
    YBL_label == "1960 to 1969" ~ 53,
    YBL_label == "1970 to 1979" ~ 43,
    YBL_label == "1980 to 1989" ~ 33,
    YBL_label == "1990 to 1999" ~ 23,
    YBL_label == "2000 to 2004" ~ 16,
    YBL_label == "2005" ~ 13,
    YBL_label == "2006" ~ 12,
    YBL_label == "2007" ~ 11,
    YBL_label == "2008" ~ 10,
    YBL_label == "2009" ~ 9,
    YBL_label == "2010" ~ 8,
    YBL_label == "2011" ~ 7,
    YBL_label == "2012" ~ 6,
    YBL_label == "2013" ~ 5,
    YBL_label == "2014" ~ 4,
    YBL_label == "2015" ~ 3,
    YBL_label == "2016" ~ 2,
    YBL_label == "2017" ~ 1,
    YBL_label == "2018" ~ 0)) %>%
select(age_struct,FHINCP,RAC1P,JWMNP, MRGP)
  
```

 
```{r}
assignment1_vars 

```


## Variables
Keeping these variables around for reference
