---
title: "Opetn Trip Planner"
author: "Sage Grey"
date: "10/2/2020"
output: html_document
---

### Libraries
```{r setup, include=FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(tidygeocoder)
library(ggmap)
```


### KML Files
```{r}
CPL_libraries <- st_read(
  "https://data.cambridgema.gov/api/geospatial/kn2z-b6eg?method=export&format=KML")

Cambridge_public_schools <- st_read("https://data.cambridgema.gov/api/geospatial/jhbq-dj88?method=export&format=KML")
```


RAN ONCE, Then deleted
opq(bbox = 'Cambridge MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/cambridge_streets.osm')


### Cambridge Streets
```{r}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

cambridge_street_features <- opq(bbox = 'Cambridge MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

cambridge_streets <- cambridge_street_features$osm_lines %>%
  st_transform(crs = MA_state_plane)


```




```{r}
ggplot(cambridge_streets) +
  geom_sf() +
  theme_map()
```


### Conntect to Open Trip Planner
**Ran Once**
path_otp <- otp_dl_jar("OTP")

```{r}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024)
otp_setup(otp = path_otp, dir = path_data, memory =1024)
otpcon <- otp_connect()
```

## CREATE SCHOOL ISOCHRONES
```{r}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = Cambridge_public_schools,  mode = "WALK", cutoffSec = 300) %>%

iso_5min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = Cambridge_public_schools, 
                mode = "CAR", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_5min_drive, iso_5min_walk)


```





## Create ISOCHROMES LIBRARIES
```{r}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = CPL_libraries, 
                mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")

iso_5min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = CPL_libraries, 
                mode = "CAR", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_5min_drive, iso_5min_walk)


```
```{r}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = CPL_libraries) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```
```{r}
ggplot(iso_all_modes) +
  geom_sf(data = cambridge_streets, color = "gray") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = CPL_libraries) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() 
```
```{r}
ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, type = "stamenwatercolor", 
                      progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = CPL_libraries) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_discrete(name = "Area that is reachable within 5 minutes",
                      labels = c("By car", "By foot"),
                      type = c("gray", "black")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")
```


#CREATE BOSTON STREETS OSM File


RAN ONCE: 
opq(bbox = 'Cambridge MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/boston_streets.osm')

```{r}
boston_street_features <- opq(bbox = 'Boston MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

boston_streets <- boston_street_features$osm_lines %>%
  st_transform(crs = MA_state_plane)


```


```{r}
ggplot(boston_streets) +
  geom_sf() +
  theme_map()
```


## ISOCHRONES FOR PUBLIC SCHOOLS
```{r}
public_schools <-st_read("http://bostonopendata-boston.opendata.arcgis.com/datasets/1d9509a8b2fd485d9ad471ba2fdb1f90_0.kml?outSR=%7B%22latestWkid%22%3A2249%2C%22wkid%22%3A102686%7D")
```


```{r}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = public_schools, 
                mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")
```



# ISOCHRONES FOR GROCCERY STORES

#ADdress of Groccery STore
```{r}
point_A <- geo(address = "761 Morton St, Mattapan, MA")
print(point_A)

```
```{r}
point_A_10min_walk <- otp_isochrone(otpcon = otpcon, fromPlace = c(-71.08928, 42.28352),
mode = "WALK", cutoffSec = 600)
```


```{r}
#Create Isochrone around Point A
point_A_10min_walk <-  otp_isochrone(otpcon = otpcon, fromPlace = c(-71.09512, 42.26820), 
                mode = "WALK", cutoffSec = 600) 
```

```{r}
#Plot Isochrone
ggplot(point_A_10min_walk) +
  annotation_map_tile(zoomin = 1, progress = "none") +
  geom_sf(fill ="blue", alpha=0.2) +
  theme_map() 


otp_stop()
```


## Attempt to use Open Street Map to Gather Groccery Data
```{r}
head(available_features())

GROCERY <- getbb("Boston") %>%
  opq() %>%
    add_osm_feature("supermarket")
```


