---
title: "Opetn Trip Planner"
author: "Sage Grey"
date: "10/10/2020"
output: 
 html_document:
    toc: yes
    toc_float: yes
    cerulean: yes
    code_folding: hide
---
# WAFFLE CONTENT
## OVERVIEW OF ASSIGNMENT

For this assignment, I decided to take a break from trying to show statistically-significant sad things about New Mexico and Mattapan, and make a very important map about the walksheds and bikesheds of places one could order waffles in Cambridge.   

## Housekeeping & Setup
### **Libraries**
```{r setup, include=FALSE}
library(osmdata)
library(opentripplanner)
library(tidyverse)
library(sf)
library(ggthemes)
library(ggspatial)
library(tidygeocoder)
library(ggmap)
library(rgeos)
```


**Attempt to Load Image**
```{{r, message = FALSE, results='hide'}
# Tried to add an image, but no dice. 
#![],("C:\Users\sageg\Desktop\svoorhees-quant\waffles_donkey.jpg")
```



**OSM Setup: Ran Once**
```{r, message = FALSE, results='hide'}
opq(bbox = 'Cambridge, MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_xml(file = 'OTP/graphs/default/cambridge_streets.osm')

```


### Load Cambridge Streets
```{r, message = FALSE, results='hide'}
MA_state_plane <- "+proj=lcc +lat_1=41.71666666666667 +lat_2=42.68333333333333 +lat_0=41 +lon_0=-71.5 +x_0=200000 +y_0=750000 +ellps=GRS80 +units=m +no_defs"

cambridge_street_features <- opq(bbox = 'Cambridge MA USA') %>%
  add_osm_feature(key = 'highway') %>%
  osmdata_sf()

cambridge_streets <- cambridge_street_features$osm_lines %>%
  st_transform(crs = MA_state_plane)


```

### Plot Cambridge Streets
```{r, message = FALSE, results='hide'}
ggplot(cambridge_streets) +
  geom_sf() +
  theme_map()
```
^ I would take big money bets that there is a tote bag with this on there somewhere in Cambridge. 

## Open Trip Planner
### Conntect to Open Trip Planner
**Ran Once**
path_otp <- otp_dl_jar("OTP")
```{r, message = FALSE, results='hide'}
path_data <- file.path(getwd(), "OTP")
path_otp <- paste(path_data, "otp.jar",sep = "/")

otp_build_graph(otp = path_otp, dir = path_data, memory = 1024)
otp_setup(otp = path_otp, dir = path_data, memory =1024)
otpcon <- otp_connect()
```



### Create Waffle Address List
```{r, message = FALSE, results='hide'}
waffle_address_list = c("1154 Massachusetts Ave, Cambridge, MA",
                        "912 Massachusetts Ave, Cambridge, MA",
"441 Cambridge St, Cambridge, MA 02141",
"3 Beacon St, Somerville, MA",
"450 Massachusetts Ave, Cambridge, MA",
"1334 Cambridge St, Cambridge, MA")
points <- geo(address = waffle_address_list, mode = "batch") 
head(points)
```

```{r, message = FALSE, results='hide'}
points <- st_as_sf(x = points,                         
           coords = c("long", "lat"),
           crs = 4326)
```

### Find Isochrones
```{r, message = FALSE, results='hide'}
waffle_points_10min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = points, 
                mode = "WALK", cutoffSec = 600) %>%
 st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")

waffle_points_10min_bike <- 
  otp_isochrone(otpcon = otpcon, fromPlace = points, 
                mode = "BICYCLE", cutoffSec = 600)%>%
 st_transform(crs = MA_state_plane) %>%
  mutate(mode = "bike")

iso_all_modes <- rbind(waffle_points_10min_walk, waffle_points_10min_bike)
```

## Walksheds of Cambridge Waffles   
Pretty close to full coverage, but definitely seeing some waffle-deserts. 
```{r}
ggplot(waffle_points_10min_walk) +
  annotation_map_tile(zoomin = 1, progress = "none") +
  geom_sf(fill ="yellowgreen", alpha=0.2) +
  theme_map() 
```


## Bikesheds of Cambridge Waffles
They did it! The free market provided waffles within biking distance of all of Cambridge!
```{r, message = FALSE, results='hide'}
ggplot(waffle_points_10min_bike) +
  annotation_map_tile(zoomin = 1, progress = "none") +
  geom_sf(fill ="sienna", alpha=0.4) +
  theme_map() 
```


## Combined! 
If you look at it right, looks like a waffle with a butter pad to me.  
Could that be accidental? Isn't it more likely part of a plan??
```{r}

right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  geom_sf(data = cambridge_streets, color = "grey") +
  geom_sf(aes(fill = mode), color =NA, alpha = 0.5) +
  geom_sf(data = points)+
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 10 minutes", labels = c("By Foot", "By Bike")) +
  theme_map() 
```

### Count the number of libraries within a 10 minute walk of Waffles


```{r, message = FALSE, results='hide'}
CPL_libraries <- st_read( 
  "https://data.cambridgema.gov/api/geospatial/kn2z-b6eg?method=export&format=KML", quiet =TRUE) %>%
  st_transform(crs = MA_state_plane) 

Cambridge_public_schools <- st_read("https://data.cambridgema.gov/api/geospatial/jhbq-dj88?method=export&format=KML", quiet =TRUE)
```


### Library Scores for our Waffle houses
Its important to know in my analysis, where one could go get waffles and then go to a public library. 
Nothing helps digestion like publicly available civic resources.  
Fortunately for us, 3 of our waffle establishment are well connected. 

```{r}
waffle_points_10min_walk <- waffle_points_10min_walk %>%
  mutate(library_score = lengths(st_covers(geometry, CPL_libraries)))
```


```{r}
ggplot(waffle_points_10min_walk) +
  annotation_map_tile(zoomin = 1, progress = "none") +
   scale_fill_distiller(palette = "Spectral")+
  geom_sf(aes(fill=library_score), alpha=.5) +
  theme_map() 
```


### Area difference of Isochrones
```{r}
iso_areas <- iso_all_modes %>%
  mutate(area = st_area(iso_all_modes)) %>%
  st_set_geometry(NULL) %>%
  pivot_wider(names_from = mode, values_from = area) 

ggplot(iso_areas, 
       aes(x = as.numeric(walk), y = as.numeric(bike))) +
  geom_point() +
  scale_x_continuous(name = 
            "Area within a ten-minute walking distance\nof a waffles\n(square km)") +
  scale_y_continuous(name = 
            "Area within a five-minute biking distance\nof a waffles\n(square km)") +
  theme_bw()
```

#### CLOSE OTP for Cambridge 
```{r}
otp_stop()
```


# NON-WAFFLE CONTENT --Not part of assignment


### Leaving this code here for my future reference, but  no need to grade beyond this point! 



## Create ISOCHROMES LIBRARIES (copied from Carole's COde)
{r, message = FALSE, results='hide'}
iso_5min_walk <- 
  otp_isochrone(otpcon = otpcon, fromPlace = CPL_libraries, 
                mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")

iso_5min_drive <- 
  otp_isochrone(otpcon = otpcon, fromPlace = CPL_libraries, 
                mode = "CAR", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes <- rbind(iso_5min_drive, iso_5min_walk)





### LOAD Cambridge KML Files
{r, message = FALSE, results='hide'}
CPL_libraries <- st_read(
  "https://data.cambridgema.gov/api/geospatial/kn2z-b6eg?method=export&format=KML")

Cambridge_public_schools <- st_read("https://data.cambridgema.gov/api/geospatial/jhbq-dj88?method=export&format=KML")



## Plot CP Library Isochrones (Carole's Code)
{r, message = FALSE, results='hide'}
right_side <- st_bbox(iso_all_modes)$xmax
left_side  <- st_bbox(iso_all_modes)$xmin
top_side <- st_bbox(iso_all_modes)$ymax
bottom_side <- st_bbox(iso_all_modes)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = CPL_libraries) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")





## CREATE SCHOOL ISOCHRONES (Modeled from Carole's COde)
{r, message = FALSE, results='hide'}
iso_5min_walk_school <- 
  otp_isochrone(otpcon = otpcon, fromPlace = Cambridge_public_schools,  mode = "WALK", cutoffSec = 300) %>%
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "walk")

iso_5min_drive_school <- 
  otp_isochrone(otpcon = otpcon, fromPlace = Cambridge_public_schools, 
                mode = "CAR", cutoffSec = 300) %>%
  
  st_transform(crs = MA_state_plane) %>%
  mutate(mode = "drive")

iso_all_modes_school <- rbind(iso_5min_drive_school, iso_5min_walk_school)



## Plot CP Library Isochrones
{r, message = FALSE, results='hide'}
right_side <- st_bbox(iso_all_modes_school)$xmax
left_side  <- st_bbox(iso_all_modes_school)$xmin
top_side <- st_bbox(iso_all_modes_school)$ymax
bottom_side <- st_bbox(iso_all_modes_school)$ymin

ggplot(iso_all_modes) +
  annotation_map_tile(zoomin = 0, progress = "none") +
  geom_sf(aes(fill = mode), alpha = 0.5) +
  geom_sf(data = CPL_libraries) +
  coord_sf(xlim = c(left_side, right_side), 
           ylim = c(bottom_side, top_side), expand = FALSE) +
  scale_fill_viridis_d(name = "Area that is reachable within 5 minutes",
                       labels = c("By car", "By foot")) +
  theme_map() +
  labs(caption = "Basemap Copyright OpenStreetMap contributors")


### Address of a groccery store
{r, message = FALSE, results='hide'}
point_A <- geo(address = "90 Hampshire St, Cambridge, MA 02139")
print(point_A)




### MAKE ISOCHRONE
{r, message = FALSE, results='hide'}
point_A_10min_walk <- otp_isochrone(otpcon = otpcon, fromPlace = c(-71.09476, 42.36874),
mode = "WALK", cutoffSec = 600)



### PLOT POINT A ISOCHRONE

{r, message = FALSE, results='hide'}
#Plot Isochrone
ggplot(point_A_10min_walk) +
  annotation_map_tile(zoomin = 1, progress = "none") +
  geom_sf(fill ="blue", alpha=0.2) +
  theme_map() 




